Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions

### 2011-08-24 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@58 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The extension of the Kolmogorov-Smirnov test to non-continuous 
null distributions is not straightforward. The formula of
the test statistic $D$ remains unchanged, but its distribution
is much more difficult to obtain; unlike the 
continuous case, it depends on the null model.
Use of the tables associated
with continuous hypothesized distributions results in conservative $p$-values
when the null distribution is discontinuous 
(see \cite{slakter1965}, \cite{goodman1954}, and \cite{massey1951}).  
In the early 1970's, 
\citet{Conover1972} developed the method implemented here
for computing $p$-values
in the case of discrete null distributions.

@@

Having obtained the test statistic, the $p$-value must then be calculated. 
For larger sample sizes, the null distribution can be approximated
with the null distribution from the classical Kolmogorov-Smirnov test. 
When an exact $p$-value is required for
smaller sample sizes, the methodology in \citet{Conover1972} is used. 
Full details of the calculations are contained in source code
of our revised function \code{ks.test()} and in Conover's paper.

@@

Functions \code{ks.test()} and \code{cvm.test()} are provided for
convenience in packages \pkg{ks.test} and \pkg{cvm.test}, respectively.
Function \code{ks.test()} offers a revision of
R's Kolmogorov-Smirnov function \code{ks.test()} from base
package \pkg{stats}, while \code{cvm.test()} is a new
function for Cram\'{e}r-von Mises tests.
Both are available from the authors
or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
\code{ks.test()} will be proposed for inclusion in \pkg{stats} in late 2010.

The revised \code{ks.test()} function supports one-sample tests for discrete
null distributions by allowing the second argument, \code{y}, to be
an empirical cumulative distribution function (an R function
with class \code{ecdf}) or an object of class \code{stepfun} specifying
a discrete distribution.  As in the original version of \code{ks.test()},
the presence of ties in the data (the first argument, \code{x}) generates a
warning unless \code{y} describes a discrete distribution.  When a discrete
distribution is specified, exact $p$-values are not available for 
two-sided alternative hypotheses, but the reported $p$-values will be
conservative; very tight bounds for interesting (small) $p$-values are
presented in \cite{Conover1972}.  For one-sided tests,
exact $p$-values are calculated using Conover's
method (when \code{exact = NULL} or \code{exact = TRUE})
if the sample size is less than or equal to 30; otherwise, asymptotic distributions
are used which are reliable in such cases.
When \code{exact = FALSE} the asymptotic distribution is used and known
to be conservative though imprecise for small samples (see \cite{Conover1972}
for details).

@@

Next, we simulate a sample of size 25 from the discrete uniform distribution on
the integers $\{1, 2, \ldots, 10\}$ and show usage of the
new \code{ks.test()} implementation.  The first is the default two-sided test,
where the reported $p$-value is a conservative upper bound for the actual
$p$-value.  In this case, the approximation may not be that tight, but this is
irrelevant for such large $p$-values. For more interesting (small) $p$-values, the
upper bound is very close to the true $p$-value (shown in \cite{Conover1972}).
\begin{Schunk}
\begin{Sinput}
> set.seed(1)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7
[14]  4  8  5  8 10  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, $p$-value = 1
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Next, we conduct the default one-sided test, where Conover's method
provides the exact $p$-value (up to the numerical precision of the
implementation):
\begin{Schunk}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

data:  rep(1, 3) 
D = 0.6667, $p$-value = 0.04938
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
If, instead, either \code{exact=FALSE} is used with the new
\code{ks.test()} function, or if the original \code{stats::ks.test()}
is used, the reported $p$-value is 0.1389 even though the test statistic
is the same.

@@

This paper presents the implementation of several nonparametric
goodness-of-fit tests for discrete null distributions.  In some cases,
the $p$-values are known to be exact.  In others, conservativeness 
and tightness of bounds have been established in the literature.  As a result,
no simulations are necessary for these methods, which were generally
developed during an era when extensive simulations may have been
prohibitively expensive or time-consuming.
However, this does raise the possibility that alternative tests relying upon
modern computational abilities could provide even greater power in certain
situations, a possible avenue for future work.

@@

\bibliography{paper}

>>>>

The extension of the Kolmogorov-Smirnov test to non-continuous 
null distributions is not straightforward. The formula of
the test statistic $D$ remains unchanged, but its distribution
is much more difficult to obtain; unlike the 
continuous case, it depends on the null model.
Use of the tables associated
with continuous hypothesized distributions results in conservative $p$-values
when the null distribution is discontinuous 
(see \cite{slakter1965}, \cite{goodman1954}, and \cite{massey1951}).  
In the early 1970's, 
\citet{Conover1972} developed the method implemented here
for computing exact one-sided $p$-values
in the case of discrete null distributions.
The method developed in \citet{gleser85} is used provide exact p-values
for two-sided tests.

@@

Having obtained the test statistic, the $p$-value must then be calculated. 
For larger sample sizes, the null distribution can be approximated
with the null distribution from the classical Kolmogorov-Smirnov test. 
When an exact $p$-value is required for
smaller sample sizes, the methodology in \citet{Conover1972} is used in
for one-sided tests.  For two-sided tests, the methods presented in
\citet{gleser85} lead to exact two-sided $p$-values.  This requires the
calculation of rectangular probabilities for uniform order statistics
as discussed by \citet{nieder81}.
Full details of the calculations are contained in source code
of our revised function \code{ks.test()} and in the papers of
Conover and Gleser.

@@

Functions \code{ks.test()} and \code{cvm.test()} are provided for
convenience in packages \pkg{ks.test} and \pkg{cvm.test}, respectively.
Function \code{ks.test()} offers a revision of
R's Kolmogorov-Smirnov function \code{ks.test()} from base
package \pkg{stats}, while \code{cvm.test()} is a new
function for Cram\'{e}r-von Mises tests.
Both are available from the authors
or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
\code{ks.test()} will be proposed for inclusion in recommended
package \pkg{stats}.

The revised \code{ks.test()} function supports one-sample tests for discrete
null distributions by allowing the second argument, \code{y}, to be
an empirical cumulative distribution function (an R function
with class \code{ecdf}) or an object of class \code{stepfun} specifying
a discrete distribution.  As in the original version of \code{ks.test()},
the presence of ties in the data (the first argument, \code{x}) generates a
warning unless \code{y} describes a discrete distribution.  
If the sample size is less than or equal to 30, exact $p$-values are provided;
otherwise, asymptotic distributions
are used which are reliable in such cases.
When \code{exact = FALSE} the asymptotic distribution is used and known
to be conservative though imprecise for small samples (see \cite{Conover1972}
for details).

@@

Next, we simulate a sample of size 25 from the discrete uniform distribution on
the integers $\{1, 2, \ldots, 10\}$ and show usage of the
new \code{ks.test()} implementation.  The first is the default two-sided test,
where the exact $p$-value is obtained using the methods of \citet{gleser85}.
\begin{Schunk}
\begin{Sinput}
> set.seed(1)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7
[14]  4  8  5  8 10  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, $p$-value = 0.9354
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Next, we conduct the default one-sided test, where Conover's method
provides the exact $p$-value (up to the numerical precision of the
implementation):
\begin{Schunk}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

data:  rep(1, 3) 
D = 0.6667, $p$-value = 0.07407
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
If, instead, either \code{exact=FALSE} is used with the new
\code{ks.test()} function, or if the original \code{stats::ks.test()}
is used, the reported $p$-value is 0.1389 even though the test statistic
is the same.

@@

This paper presents the implementation of several nonparametric
goodness-of-fit tests for discrete null distributions.  
In some cases the $p$-values are known to be exact.  In others,
conservativeness in special cases with small $p$-values
has been established.  As a result,
no simulations are necessary for these methods, which were generally
developed during an era when extensive simulations may have been
prohibitively expensive or time-consuming.
However, this does raise the possibility that alternative tests relying upon
modern computational abilities could provide even greater power in certain
situations, a possible avenue for future work.

@@

\section{Acknowledgement}

We are grateful to Le-Minh Ho for his implementation of the algorithm
described in \citet{nieder81} for calculating rectangular probabilities
of uniform order statistics.

\bibliography{ArnoldEmerson}
